\mytitle{}

### 2018-10-19 paper/paper.tex
Paperial Evolution (?)

\documentclass[conference, letterpaper, 10pt, times]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}

@@

%bib
\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,mincrossrefs=5,sortcites,backend=bibtex,uniquelist=false]{biblatex}
\addbibresource{papers-off.bib}
\addbibresource{confs-off.bib}
\addbibresource{books-off.bib}
\addbibresource{rfc.bib}
\addbibresource{misc.bib}

@@

@@

?? Reference to the possible mistreatment of RL in our domain? The problem of ``blind application'' so hated by \textcite{DBLP:conf/sp/SommerP10}.

This paper contributes:
\begin{itemize}
	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A flow-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

\section{Background}
%?? Introduce RL, related definitions etc.
\subsection{Reinforcement Learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov decision process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.

@@

There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
This transform, as an example of linear function approximation, then necessitates the use of \emph{1-step semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ taking $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$ by continually updating $\wvec{t}$ as follows:
\begin{subequations}
	\begin{gather}
	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
	\end{gather}
	\label{eqn:sg-sarsa}
	taking $\nabla$ with respect to $\wvec{}$.
\end{subequations}

@@

DDoS attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
Hosts often participate unwillingly, often having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most vulnerable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed to appear as though the originated from the victim, significantly increasing a botnet's throughput while masking the identity of each participant.
\emph{Transit-link attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete benefits associated with this problem.
We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.

@@

\section{Threat Model}

@@

We assume that attacks are flooding-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.
Finally, we do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.

\section{Environment and RL Algorithm}\label{sec:environment-and-rl-algorithm}

?? Probably rethink the title of this section.

%?? State the problem and what we want to achieve.
To best discover how the use of RL techniques in the field of network intrusion prevention may yet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback loop-like control of the network to detect and prevent DDoS attacks.
We begin with a reimplementation and more principled re-examination of their work to further develop the field.

@@

The reward functions as specified here were introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
\begin{subequations}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
\begin{gather}
\cond{f} = [\load{s} > U_s],\\
\cond{c}{(v)} = [\load{l_v} > U_s/k],\\
R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\\
R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},
\end{gather}
where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
\end{subequations}

@@

?? TODO: Make repo public...

\subsection{Service Guarantees and Host Density}
The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions as defined become less granular.

@@

%\Cref{fig:mplex-tcp} shows that this trend is present.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
\begin{equation}
\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
\end{equation}
for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
\Cref{fig:nginx-coffin-nail} shows that this trend is present.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.
As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP.

@@

Measurements during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
This is reassuring when measured alongside the insights from other work.
\textcite{DBLP:conf/sigcomm/ChenL0L18} observed that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
%Most flows are short, and flow-size follows a heavy-tailed distribution.
That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used to maintain consistency over traffic conditions.
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate traffic the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
This provides two modes of operation:
\begin{itemize}
	\item \emph{long flow} testing has each host constantly request a large binary, such as a tarball of GCC8;
	\item \emph{request-chain} testing has each host request a random file known to exist within the website, followed by any dependent resources (stylesheets, images, etc.). This presents a more balanced distribution of flow duration and size.
\end{itemize}
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
Accordingly, reward functions aim to maximise the downstream throughput.

@@

@@

?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)?

@@

?? What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge? I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}. We can extend the team-based formulations similarly. This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server. The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other!

@@

?? Other recent work in the prevention of DDoS? (i.e., non-RL).
\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} had it as part of their evaluation, BGP-like rerouting techniques (specifically for transit-link attacks, but also for `normal' attacks) \cite{RoutingAroundCongestion}

%?? Abuses of RL 
Earnest, well-considered application of RL towards the challenge of intrusion detection/prevention has seen comparatively little examination.
Past work exhibits treatment of the paradigm as a traditional classifier for anomaly detection \cite{shamshirband2014anomaly} and DDoS prevention \cite{DBLP:conf/mates/ServinK08}.
Given that one of the main strengths of RL techniques is their ability to control ongoing interaction and to adapt by observing the concrete effects of actions taken, such works fail to see how best to apply the rich literature on the subject.

For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
?? Good Direct-control work? What this work is based on (which directly captures the spirit of RL) \cite{DBLP:phd/ethos/Malialis14, DBLP:journals/eaai/MalialisK15}

>>>>

% !TeX document-id = {e7fef4b9-afbc-4e81-94aa-c4847183d047}
% !BIB program = biber
\documentclass[conference, letterpaper, 10pt, times]{IEEEtran}

%\usepackage[T1]{fontenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}

@@

%bib
\usepackage[style=ieee,maxnames=3,mincitenames=1,maxcitenames=2,maxbibnames=99,
mincrossrefs=99,minxrefs=99,
sortcites,
%backend=bibtex,
uniquelist=false]{biblatex}
\addbibresource{papers-off.bib}
\addbibresource{confs-off.bib}
\addbibresource{books-off.bib}
\addbibresource{rfc.bib}
\addbibresource{misc.bib}

@@

\newcommand{\fakepara}[1]{\noindent\textbf{#1:}}

@@

To date, there have been few applications of this class of algorithms towards intrusion prevention which make use of their full potential for online control, rather than using them as the basis of a classifier.
We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
In service of this goal, this paper contributes:
\begin{itemize}
	\item A DDoS mitigation system based on direct-control reinforcement learning designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:rethinking-the-state-space}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}---a deconstruction and empirical study on their formulation, flaws, and risk factors concerning many traffic distributions (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics which packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item A flow-level granularity approach to RL-driven DDoS prevention, improving upon past aggregate-based approaches (\cref{sec:rethinking-the-state-space}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

\section{Background and Threat Model}
%?? Introduce RL, related definitions etc.
\subsection{Reinforcement Learning}
\emph{Reinforcement learning} (RL) is a variant of machine learning principally concerned with with training an agent to choose an optimal sequence of actions in the pursuit of a given task \cite{RL2E}.
We assume the agent has a certain amount of knowledge whenever a decision must be made: at any point in time $t$ it knows which \emph{state} it is in ($S_t \in \mathcal{S}$), the set of \emph{actions} which are available to it ($\operatorname{A}(S_t) \subseteq \mathcal{A}$) and, if available, a numeric \emph{reward} obtained from the last action chosen ($R_t \in \mathbb{R}, A_{t-1} \in \operatorname{A}(S_{t-1})$).
This model of system interaction is best described as a \emph{Markov decision process} (MDP).
RL methods combine this information with a current \emph{policy} $\pi$ to determine which action should be taken: such a choice need not be optimal, if an agent needs to further explore some region of the state space.
The policy is then further refined by updating value estimates for state-action pairs or via policy gradient methods, meaning that RL-based approaches learn adaptively and online if reward functions are available in the environment they are deployed in.
From any point in a sequence of decisions, we may describe the sum of rewards yet to come as the \emph{discounted return}, $G_t = R_{t+1} + \gamma R_{t+2} + \gamma^2 R_{t+2} + \ldots$, choosing the discount factor $\gamma \in [0,1]$ to determine how crucial future rewards are vis-\`{a}-vis the current state.
Formally, an agent's goal is to choose actions which maximise the \emph{expected discounted return} $\operatorname{\mathbb{E}_{\pi}}[G_t]$.

@@

There is immense variation in \emph{how} policies and/or values may be learned, reliant upon the learning environment, problem and required convergence guarantees.
In particular, we shall focus on methods which choose actions according to their value estimates from the current state.
To cope with a continuous state- and/or action-space, one valuable technique is to employ \emph{tile-coding} \cite[pp.\ \numrange{217}{221}]{RL2E}---this converts our state from $\mathbb{R}^d \times \mathcal{A}$ into a sparse boolean feature vector $\operatorname{\mathbf{x}}(s, a)$ by subdividing it into a number of overlapping $(d+ \dim{\mathcal{A}})$-dimensional grids with an optional bias component.
The number of tilings has a direct effect on the resolution of the discretisation.
Entries of $\operatorname{\mathbf{x}}(s, a)$ are set to 1 if the state-action pair exists within the corresponding tile.
This transform, as an example of linear function approximation, then necessitates the use of \emph{1-step semi-gradient Sarsa} \cite[pp.\ \numrange{243}{244}]{RL2E}.
Given a weight vector $\wvec{0}=\bm{0}$ and learning rate $\alpha$, we may approximate an action $a$'s value in any state $s$ taking $\operatorname{q}(s, a) \approx \acval{s}{a}{\wvec{}}$ by continually updating $\wvec{t}$ as follows:
\begin{subequations}
	\begin{gather}
	\acval{s}{a}{\wvec{}} = \wvec{}^{\top} \operatorname{\mathbf{x}}(s, a),\\
	\delta_t = R_{t+1} + \gamma \acval{S_{t+1}}{A_{t+1}}{\wvec{t}} - \acval{S_t}{A_t}{\wvec{t}},\\
	\bm{w}_{t+1} = \bm{w}_{t} + \alpha \delta_t \nabla{\acval{S_t}{A_t}{\wvec{t}}},
	\end{gather}
	\label{eqn:sg-sarsa}
	taking $\nabla$ with respect to $\wvec{}$.
\end{subequations}

@@

DDoS attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this either by exploiting peculiarities of operating system or application behaviour (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{flooding-based attacks}).
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

Although there are variations of each class of attack, flooding attacks are the most relevant to our work.
\emph{Amplification attacks} exploit the presence of services who eagerly send large replies in response to small requests, where UDP-based services like DNS and NTP are most vulnerable \cite{DBLP:conf/ndss/Rossow14, DBLP:conf/uss/KuhrerHRH14}.
Malicious hosts send many small requests, spoofed to appear as though the originated from the victim, significantly increasing a botnet's throughput while masking the identity of each participant.
\emph{Transit-link/link-flooding attacks} have been the subject of recent attention, wherein malicious traffic is forwarded across core links needed to reach a target (but not to the target itself) \cite{DBLP:conf/sp/KangLG13, DBLP:conf/esorics/StuderP09}.

\subsection{Motivation}
%?? What makes RL a suitable method for network anomaly detection, what features are most relevant?
%?? Point I was thinking of: feedback-loop-like model allows monitoring \emph{after} an action is taken to (in theory) allow forgiveness of mistakenly punished flows. This does hinge on taking a flow-by-flow look at the state space, but if we can combine knowledge of current state (duh!), the last action taken (i.e. an indicator of our previous assessment [such as high pdrop $\implies$ bad]) then perhaps a flow which falls off identically to a legit flow can be rescued.
Moving beyond the overt benefits of choosing RL-based defences for coping with non-stationary problems, we believe that there are concrete reasons for their use here.
We have seen that for other domains in particular, misclassification is a serious problem, which can introduce \emph{collateral damage} in the context of DDoS prevention.
In theory, the feedback-loop-like model allows us to monitor flows \emph{after} an action is taken to allow forgiveness of mistakenly punished flows.
This does rely upon the ability to take a flow-by-flow view of the state space, but if we can combine knowledge of current state with the last applied action, then perhaps a flow which falls off identically to a legitimate flow can be rescued.

@@

\subsection{Threat Model}

@@

?? Oultine attacker's goals, particularly wrt SPIFFY.
?? They want to impact fair-share bandwidth allocation to flows (minimise), they have no worry about being unmasked (there are many available reflectors, which don't need to be bought etc...) [cite the rossow papers about amplification hell].
We assume that attacks are bandwidth-based DDoS attacks with the structure of an \emph{amplification attack}, and that traffic aggregates at the target (unlike in a \emph{transit-link attack}).
The addresses of any machines taking part in an attack are not revealed to the target, save for the fairly stable set of unwitting reflector nodes.

Finally, we do not assume that an attacker has white-box access to the parameters underlying an agent's policy, or that they will attempt to intelligently modify flow/system state to indirectly control an agent \cite{DBLP:conf/eurosp/PapernotMJFCS16, DBLP:conf/eurosp/PapernotMSW18, DBLP:journals/corr/HuangPGDA17, DBLP:conf/sp/Carlini017}.
While they may be able to perform some degree of reverse engineering by observing the health of their own legitimate canary flows, ``stealing'' the policy through observation \cite{DBLP:conf/uss/TramerZJRR16}, investigating whether perturbations would persist in a medium as volatile as network traffic statistics falls outside of the scope of this work.
The same observation extends to the possibility of poisoning attacks \cite{DBLP:journals/jmlr/KloftL10, DBLP:conf/acsac/ShenTS16}; to the best of our knowledge, no studies have been undertaken surrounding the mis-training of RL agents.

\section{Analysing Existing Work}\label{sec:environment-and-rl-algorithm}
%?? State the problem and what we want to achieve.
To best discover how the use of RL techniques in the field of network intrusion prevention may yet be improved, it's prudent that we start by looking for weaknesses in work which exists today.
The closest approach which is available within this field is that of \textcite{DBLP:journals/eaai/MalialisK15}, and their contribution in applying RL to the task of intrusion prevention is significant: their work helped to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
We begin with a reimplementation and more principled re-examination of their work to further develop the state-of-the-art.

@@

We employ the reward functions introduced by \textcite{DBLP:journals/eaai/MalialisK15}; in particular, we make use of their global reward as a performance metric and their \emph{coordinated team learning} (CTL) reward for training.
These assume the presence of a classifier, $g(\cdot)$, which estimates the volume of `legitimate traffic' received at any point in the network.
At time $t$ an agent $v$ (with team leader $l_v$) receives either $R_{t}^{\mathit{Global}}$ or $R_{t}^{\mathit{CTL}}$:
\begin{subequations}
\newcommand{\cond}[2]{\operatorname{c}_{#1,t}#2}
\newcommand{\load}[1]{\operatorname{load}_{t}(#1)}
\begin{gather}
\cond{f} = [\load{s} > U_s],\\
\cond{c}{(v)} = [\load{l_v} > U_s/k],\\
R_{t}^{\mathit{Global}} = (1 - \cond{f}) \frac{g(\load{s})}{g(\load{\mathit{total}})} - \cond{f},\\
R_{t}^{\mathit{CTL}}(v) = (1 - \cond{f}\cond{c}{(v)}) \frac{g(\load{l_v})}{g(\load{\mathit{team}})} \allowbreak - \cond{f} \cond{c}{(v)},
\end{gather}
where $\load{\mathit{total}}$ and $\load{\mathit{team}}$ refer to the load across all egress points in the network and in $v$'s team, respectively.
\end{subequations}

@@

%?? TODO: Make repo public...

\subsection{Service Guarantees and Host Density}
The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions (as defined above) become less granular.

@@

%\Cref{fig:mplex-tcp} shows that this trend is present.
Given the UDP-like behaviour of Tcpreplay, being able to replicate the online performance of \citeauthor{DBLP:journals/eaai/MalialisK15} in the $n=2$ case suggests that the numeric simulation their work is entirely based upon does not capture the key interactions between TCP flows and packet drop as a control action.
From the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97}, through which the performance characteristics of TCP under packet drop are well-known and well-understood, their results should be considerably worse in practice:
\begin{equation}
\operatorname{BW} = \frac{\operatorname{MSS}}{\operatorname{RTT}} \frac{C}{\sqrt{p}},
\end{equation}
for the \emph{maximum segment size} (MSS), \emph{round trip time} (RTT), a choice of constant $C \approx{} \sqrt{3/2}$ dependent on system assumptions, and the probability $p \in (0, 1]$ that a packet is lost.
Accordingly, benign TCP flows are severely harmed by \emph{any} packet drop, so collateral damage to these hosts is more significant.
Good-faith TCP congestion avoidance causes these flows to attempt to back off very quickly.
\Cref{fig:nginx-coffin-nail} shows that this trend is present.
%?? Talk about the same major points you made in the presentation here (i.e. TCP interactions) w/ pushback. Or in the evaluation...
The TCP traffic as it appears here is generated according to \Cref{sec:a-new-normal}.
UDP, however, behaves as expected---malicious hosts are expected to exhibit the same behaviour, given that congestion is their primary goal.

These results are of particular importance due to TCP's prevalence within the Internet: it carries \SIrange{79}{94}{\percent} of packets, corresponding to \SIrange{89}{98}{\percent} of data transferred by volume \cite{DBLP:conf/saint/ZhangDJC09}.
As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP.

@@

Overall, each episode takes around \SI{10}{\minute} to run, while each set of \num{10} requires around \SI{2}{\hour} due to additional set-up/tear-down costs associated with mininet.
Measurements taken during each of these experiments indicated that the cost of computing any action is typically within \SIrange{80}{100}{\micro\second}.
This is reassuring when measured alongside the insights from other work.
\Textcite{DBLP:conf/sigcomm/ChenL0L18} observe that, ideally, actions must be computed and taken within \SI{1}{\milli\second} to have a meaningful affect on short flows.
%Most flows are short, and flow-size follows a heavy-tailed distribution.
That our starting point falls significantly below this threshold allows us to safely consider more costly actions or larger state spaces, which would typically increase the computational cost.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate traffic the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
This provides two modes of operation:
\begin{itemize}
	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
	\item \emph{request-chain} testing has each host request a random file known to exist within the website, followed by any dependent resources (stylesheets, images, etc.). This presents a more balanced distribution of flow duration and size.
\end{itemize}
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
Accordingly, reward functions aim to maximise the downstream throughput.

@@

?? note---probably put in the experiments that Simon suggested at this point as justification for why we chose what we did...

@@

?? What concessions will we have to make in order to make per-flow processing more viable? Intelligent sampling/reanalysis of flows when needed (i.e. an external heuristic guiding method)? In SPIFFY's \cite{DBLP:conf/ndss/KangGS16} evaluation, we see clearly that it can take around \SI{2}{\second} for a flow to react fully to a rate increase---I think for the TCP step it may be wise to factor this in, too!

@@

?? What other consequences are there? We're training each agent on all the flows it's acting on which means slight theoretical changes (we have multiple experience traces, and each uses its last state + new state + new target value (reward) as an update to the \emph{current model}).

?? What might we do for a reward function in the absence of heuristic estimates and/or explicit a priori knowledge? I think a good candidate is the sum of up and down throughputs (normalised by capacity sum), so long as \emph{neither exceeds the link capacity}. We can extend the team-based formulations similarly. This, in theory, promotes traffic diversity since it's not like flooding-based DDoS attacks are going to submit meaningful work to a server. The intuition, I suppose, is that certain classes of flow will have a small footprint in one direction which causes a sizeable increase in the other! Alternatively, monitor the health of canary flows which cross the team boundary (i.e. only one in-out link).

@@

%?? Try and compare my work here when possible?

\fakepara{DDoS Prevention}
%?? Other recent work in the prevention of DDoS? (i.e., non-RL).
\Textcite{DBLP:conf/lcn/BragaMP10} have examined the detection of ongoing (flooding-based) DDoS attacks through \emph{self-organising maps}, making use of SDN to gather statistics effectively.
Many of their features aren't overly relevant, as their focus is not active defence or discovering \emph{which} hosts are contributing to an attack.
%?? What does it do? detects state, not which flows are contributing...

\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.

\emph{Athena} \cite{DBLP:conf/dsn/LeeKSPY17} is a more generalised SDN framework for intrusion detection, but has shown the use of a \emph{k nearest-neighbours} classifier to detect individual attack flows.
Although heavyweight (and proven to be effective compared with \textcite{DBLP:conf/lcn/BragaMP10}), their comparison against SPIFFY lacks the quantitative evidence required to understand how the system compares.

\Textcite{DBLP:conf/sp/SmithS18} present techniques based on AS-level routing to tackle both transit-link and flooding-based attacks.
This view is taken due to the perceived cost of per-stream classification and inherent sensitivity to adversarial examples or crafted input.
The approach is creative, relying upon BGP \emph{fraudulent route reverse poisoning} to preserve traffic to a target AS, but unlike SPIFFY the approach doesn't actually \emph{remove} the congestion.
Because of this, traditional flooding-based attacks aren't fully alleviated.
